Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-11-17 manuscript/reviewer_reply.Rmd

Table 1 caption: without getting to more methods (below), I don’t understand the sentence ‘Environmental models were trained using all sites together, without information regarding which transects occurred at which site or during which year’.

Table 1: Just my personal preference, but I’d include the check marks and replace the ‘Xs’ with blank cells

Line 194-221: It seems for each approach you only considered the full models with all environmental covariates? I’m curious whether any of the covariates appeared to not improve explanatory power, and whether you considered models that dropped these uninformative predictors?

> "Line 216: Cite randomForest / caret/ party R package – or whatever else you used?"

**The randomForest package was cited in the final methods paragraph. We've cited it where it is first mentioned as well. **

@@

see general comments
Comments for the author

In this paper, the authors use the incredibly rich data from the Breeding Bird Survey to examine how well various models perform at predicting bird richness at sites throughout North America. They train models on data from 1982-2003 and attempt to predict data from 2004-2013. They examine the performance of six different model types, two of which are null or baseline models.

I find it a bit surprising that the authors think that any of these richness models could predict a high level of accuracy for theses ten years. I think of richness estimations as a very coarse technique for looking at broad shifts in richness over relatively long time periods rather than these 5-species (~5-7%) shifts in richness on annual bases indicated in Figure 2. The authors don’t provide a lot of their raw data on sites other than as examples, but my impression is that they are pursuing an unrealistic level of accuracy and detail. That being said, given what the paper does, it’s okay. The models seem to be performed well. The authors give all of the keywords to indicate that they know what they are doing.

Here are some comments:
- Birds are not the best system for doing any sort of distribution analyses. They are incredibly vagile, hard to detect, and the distributions of breeding birds are affected by many components besides environmental variables (particularly interspecific interactions). No assessment of detectability is included here other than accounting for observer identity. However, I do realize that BBS data are detailed and compelling data to use given what a large amount of detailed time-series it provides.
- Abstract: The authors forecast the data and then test it using known data. Hindcast only refers to back-in-time predictions.

The definition of hindcasting varies by field. We use it here as it's used in meteorology, where any prediction of observations already available is generally a hindcast, and forecasts are only for predictions which are actually in the future. See Jolliffe & Stephenson 2003 (in manuscript) section 1.4.2 for a longer discussion. We've clarified this in the text.

- 25-29: rather than relay the type of information that the paper will provide, the authors should summarize their results and findings
- 57-58: this is not the definition of hindcasting. This is the definition of testing. Also, the authors exclude the many deeper-time hindcasting studies that have been performed on these types of models

See comment above on hindcasting.

- Methods: the authors perform their analyses on data that are collected from an incredibly dynamic time period. I wouldn’t expect species to have settled into a richness pattern given the incredibly disrupted nature of the last 32 years. Some strategies to try to handle these issues might be to train and predict at different temporal scales and to both forecast and hindcast the predictions. (e.g. train on 2004-2013 and try to predict 1982-1992 data + 1993-2003 data). If the authors were to make splits based on expected periods of dynamism, they could better tell whether the drivers that are included in the models are accurate predictors.
- 124-130: I didn’t see where future projections were performed/used in this paper. Why is this here?
- Fig. 2: Panels not labeled. No key for colors. Somewhat confusing.
- Fig. 6 caption: “compoonents”
- 466: only other mention of future forecasts… not sure why this is included in this paper.

### 2017-11-17 manuscript/draft.Rmd

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. The BBS
data was filtered to exclude all nocturnal, cepuscular, and aquatic species
[since these species are not well sampled by BBS methods; @hurlbert2005], as
well as unidentified species, and hybrids. All data from surveys that did not
meet BBS quality criteria were also excluded.

>>>>

**Richness data.** Bird species richness was obtained from the North American
Breeding Bird Survey (BBS) [@pardieck2017] using the Data Retriever Python
package [@morris2013] and rdataretriever R package [@rdataretriever]. BBS
observations are three-minute point counts made at 50 fixed locations along a
40km route. Here we denote each route as a site and summarize richness as the
total species observed at all 50 locations in each surveyed year. Prior to
summarizing the data was filtered to exclude all nocturnal, cepuscular, and
aquatic species [since these species are not well sampled by BBS methods;
@hurlbert2005], as well as unidentified species, and hybrids. All data from
surveys that did not meet BBS quality criteria were also excluded.
